---
title: "Initial Analysis"
output: html_document
date: "2024-11-25"
---

## This file includes all the data cleaning  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(dplyr)
```


## FOIA DATASET CLEANING

```{r foia data read in, echo = TRUE}
foia_data = read.csv("foia-504-fy2010-present-asof-240930.csv")
glimpse(foia_data)
```

The *As of Date* is the date that all of the data was recorded we are going to 
remove this row because it has no use. There is only one unique date.

```{r unique dates in foia , echo = TRUE}
unique(foia_data$AsOfDate)
foia_data = foia_data %>% select(-AsOfDate)
```

All business in the dataset are *504 small business* so this column can be removed
as well.

```{r unique programs in foia, echo = TRUE}
unique(foia_data$Program)
foia_data = foia_data %>% select(-Program)
```

```{r, echo = TRUE}
head(foia_data)
```
First we are going to take a look at the lender variables which are BorrName,
BorrStreet..., CDC_Name, CDC_Street.... , ThridPartyLender_Name....

Checking for missing values in the foia dataset but focusing in on the lender
information

```{r summarize NAs in foia, echo = TRUE}
foia_data %>% summarize_all(~ sum(is.na(.) | . == ""))
```
Majority of the observations that have missing values for Third Party Lender
have no missing values for CDC. This checks out as this dataset is from the
official SBA page.

We are going to get rid of the values that have no information on the name,
city and state.

```{r filter NA for city and state FOIA, echo = TRUE}


foia_data = foia_data %>% filter(!(ThirdPartyLender_Name == "" &
                              ThirdPartyLender_City == "" &
                              ThirdPartyLender_State == ""))

foia_data = foia_data %>% filter(!(CDC_Name == ""))

```

Values with information on the Name can be found using another
dataset https://www.fdic.gov/ - includes information on all FDIC supported banks
including their name, city and state among other things..
The locations dataset has location information from all public banks.
We are going to utilize this dataset to find missing city and state information.

The name data is in title case so we are going to change the thirdpartylenders
names to title case.

We are also going to rename the lender_location NAME column to ThirdPartyLender_Name
so that we can do the left join.

```{r lender location join, echo = TRUE}
lender_location = read_csv("locations.csv", show_col_types = FALSE)
head(lender_location, n = 3)
foia_data = foia_data %>% mutate(ThirdPartyLender_Name = 
                              str_to_title(foia_data$ThirdPartyLender_Name))
lender_location = lender_location %>% rename(ThirdPartyLender_Name = "NAME")

```

Trying to resolve duplicates in both datasets and duplicate lender_location.
We want to make sure there are only unique values.
Then, performing a left join to merge city and state information into 
foia_data_lender based on ThirdPartyLender_Name.

```{r foia and lender location duplicates, echo = TRUE}
foia_duplicates <- foia_data %>%
  count(ThirdPartyLender_Name) %>%
  filter(n > 1)
head(foia_duplicates,5)

location_duplicates <- lender_location %>%
  count(ThirdPartyLender_Name) %>%
  filter(n > 1)


lender_location <- lender_location %>%
  group_by(ThirdPartyLender_Name) %>%
  slice(1) %>%
  ungroup()

foia_data <- foia_data %>%
  left_join(lender_location, by = "ThirdPartyLender_Name") %>%
  select(names(foia_data)) #selecting only the columns from foia_data_lender
#the other dataset doesn't matter

foia_data <- foia_data %>%
  mutate(
    ThirdPartyLender_City = if_else(
      is.na(ThirdPartyLender_City) | ThirdPartyLender_City == "",
      "Unknown",  
      ThirdPartyLender_City
    ),
    ThirdPartyLender_State = if_else(
      is.na(ThirdPartyLender_State) | ThirdPartyLender_State == "",
      "Unknown", 
      ThirdPartyLender_State
    )
  )
```
We reduced the amount of missing values from 192 and 197 to 22 and 27.
We are just going to omit the rest of the missing values due to time. But 
some strategies to figure out the rest of the missing values would be:

1. look up manually the city and state based on the thirdpartylender name 
2. Find another dataset that includes other banks.

There are 106583 observations, so omitting 27 observations won't make any 
significant changes to our analysis.
We are now left with 106556 observations.

```{r foia NA check third party lender, echo = TRUE}
foia_data %>% summarize_all(~ sum(is.na(.) | . == "" | . == "Unknown"))
foia_data = foia_data %>% filter(!(ThirdPartyLender_State == "" | 
                              ThirdPartyLender_State == "Unknown" |
                              ThirdPartyLender_City == ""| 
                              ThirdPartyLender_City == "Unknown"))
foia_data %>% summarize_all(~ sum(is.na(.) | . == "" | . == "Unknown"))
dim(foia_data)
```

We have now successfully cleaned the lender portion of the data 
Doing a final glimpse check to see if all the variables make sense.

```{r final check of foia data, echo = TRUE}
glimpse(foia_data)
```

Looking at this we see some inconsistencies with the casing.
e.g. Mint Hill but New york. 

There are some words in parentheses but we are going to keep everything and 
just change the casing. If we need to focus on a specific city in analysis 
we can just extract the city using a str_detect and regex.

We ended up changing the Borrows Name, street and city to title case. 

```{r str change BORR CITY foia , echo = TRUE}
head(unique(foia_data$BorrCity),10)

foia_data = foia_data %>%
  mutate(BorrCity = str_to_title(foia_data$BorrCity),
         BorrStreet = str_to_title(foia_data$BorrStreet),
         BorrName = str_to_title(foia_data$BorrName))


```


## Loan information and industry part of the dataset (recipients of the loan)

First look into missing values in the foia dataset.
Here we can see that ChargeOffDate needs to be investigated

```{r summarize NA foia data, echo = FALSE}
foia_data %>% summarize_all(~ sum(is.na(.)))
```


Previously, we noticed that ChargeOffDate has a lot of NA's, so let's try to 
clean that up.

```{r charge off date removal, echo = TRUE}
foia_data = foia_data %>%
  mutate(ChargeOffDate = ifelse(is.na(ChargeOffDate), "Not Applicable", as.character(ChargeOffDate))) 
```

**FranchiseName** and **PaidInFullDate** have a lot of empty rows, 
so let's replace those values with a placeholder to make it easier to 
analyze the data.
For missing values in CongressionalDistrict, we will replace them with 'Unknown.'

```{r foia " " handling, echo = TRUE}
foia_data = foia_data %>%
  mutate(
    FranchiseName = ifelse(FranchiseName == "", "Not Applicable", FranchiseName),
    PaidInFullDate = ifelse(PaidInFullDate == "", "Not Applicable", PaidInFullDate),
    CongressionalDistrict = ifelse(is.na(CongressionalDistrict), "Unknown", 
                                   CongressionalDistrict)
  )
```

Other than NAs, many variables have empty values as well that we need to 
clean up. 

```{r checking empty " ", echo = TRUE}
foia_data %>% summarize_all(~ sum(is.na(.) | . == ""))
```

Unlike with FranchiseName and PaidInFullDate which specify in the description 
of the variable that they apply only if applicable, these variables with empty 
values above do not do that. But we do not want to simply remove these rows, 
so we replace the empty values with 'Unknown' to remain consistent.

```{r, echo = TRUE}
foia_data <- foia_data %>%
  mutate(
    FirstDisbursementDate = ifelse(FirstDisbursementDate == "", "Unknown", 
                                   FirstDisbursementDate),
    NaicsDescription = ifelse(NaicsDescription == "", "Unknown", NaicsDescription),
    FranchiseCode = ifelse(FranchiseCode == "", "Unknown", FranchiseCode),
    ProjectCounty = ifelse(ProjectCounty == "", "Unknown", ProjectCounty),
    CongressionalDistrict = ifelse(CongressionalDistrict == "", "Unknown", CongressionalDistrict),
    BusinessAge = ifelse(BusinessAge == "", "Unknown", BusinessAge),
    LoanStatus = ifelse(LoanStatus == "", "Unknown", LoanStatus)
  )
```

Last thing we are going to do is make a column to match each particular industry
by their NAICS code . Although there is a column called "NAICS Description" and 
"BusinessType" these aren't descriptive at all and just give either the names of 
the company (which isn't helpful for us), or a broad definition.

- For NAICS classification we are only going to focus on the first two digits
which gives us information on the sector. Not the specific sub-sector or industry
group. https://www.naics.com/search/ is where I am getting information from 
for this particular block of code.


(for my first time around I got about 60,000 unknown observations
then I realized that I forgot to do case_when statements for the right table 
of industries on the website......)

```{r, echo = TRUE}

foia_data = foia_data %>% mutate(NaicsCode = as.character(NaicsCode)
) %>% mutate(NaicsIndustry = case_when(
  str_detect(NaicsCode, "^11\\d{4}$") ~ "Agriculture",
  str_detect(NaicsCode, "^21\\d{4}$") ~ "Mining",
  str_detect(NaicsCode, "^22\\d{4}$") ~ "Utilities",
  str_detect(NaicsCode, "^23\\d{4}$") ~ "Construction",
  str_detect(NaicsCode, "^3[1-3]\\d{4}$") ~ "Manufacturing",
  str_detect(NaicsCode, "^42\\d{4}$") ~ "Wholesale Trade",
  str_detect(NaicsCode, "^4[4-5]\\d{4}$")  ~ "Retail Trade",
  str_detect(NaicsCode, "^4[8-9]\\d{4}$") ~ "Transportation and Warehousing",
  str_detect(NaicsCode, "^51\\d{4}$")  ~ "Information",
  str_detect(NaicsCode, "^52\\d{4}$") ~ "Finance and Insurance",
  str_detect(NaicsCode, "^53\\d{4}$") ~ "Real Estate",
  str_detect(NaicsCode, "^54\\d{4}$") ~ "Professional Services",
  str_detect(NaicsCode, "^55\\d{4}$") ~ "Management of Companies",
  str_detect(NaicsCode, "^56\\d{4}$") ~ "Administrative and Support Waste Services",
  str_detect(NaicsCode, "^61\\d{4}$") ~ "Educational Services",
  str_detect(NaicsCode, "^62\\d{4}$") ~ "Health Care and Social Assistance",
  str_detect(NaicsCode, "^71\\d{4}$") ~ "Arts, Entertainment, and Recreation",
  str_detect(NaicsCode, "^72\\d{4}$") ~ "Accommodation and Food services",
  str_detect(NaicsCode, "^81\\d{4}$") ~ "Other services (except Public Adminstration)",
  str_detect(NaicsCode, "^92\\d{4}$") ~ "Public Administration",
  TRUE ~ "Unknown"
))

```


Interesting enough we have some Naics Codes that have no description,
no industry about 300 in total (Not printing them all).
They all have the same NaicsCode of 0 which our assumption is the equivalence of 
an NA value.

```{r, echo = TRUE}
foia_data %>% select(NaicsCode, NaicsDescription, NaicsIndustry) %>%
  filter(NaicsIndustry == "Unknown") %>% slice_head(n = 5)
```

We are going to do some investigation into these missing values and see if there
are any discrepancies.

1. we are going to get rid of all observations where the franchiseCode and 
Franchise Name are unavailable because there would be no way to find the 
companies NAICS code (this step got rid of about 250 observations)
2. We could find the NAICS code through an API but on search you have to pay for
the service. So we are just going to omit these values.
3. You could also manually look it up through the same website
https://opencorporates.com/ but that would be tedious.

```{r filter out unkowns for both naics and franchisecode, echo = TRUE}
foia_data %>% filter(NaicsIndustry == "Unknown") %>% head(n = 5)

foia_data = 
  foia_data %>% filter(!(NaicsIndustry == "Unknown" & FranchiseCode == "Unknown"))

foia_data = foia_data %>% filter(!(NaicsIndustry == "Unknown"))

```

Just a quick first glance we can see there are a lot of industries in this 
dataset which is great!

```{r industry in foia, echo = TRUE}
foia_data %>% group_by(NaicsIndustry) %>% summarize(count = n())
```

Saving the new cleaned file as a csv.
```{r foia cleaned file, echo = TRUE}
write_csv(foia_data, "foia_data.csv")
```




## ASM DATA CLEANING

```{r asm intial glimpse, echo = TRUE}
ASM_data = read.csv("ASMAREA2017.AM1831BASIC03-2024-11-22T003727.csv")
glimpse(ASM_data)
```

Let's check for any NAs in this dataset. 
```{r asm na check, echo = TRUE}
ASM_data %>%
  summarize_all(~ sum(is.na(.)))
```

There are no NAs present, but the column names are too long and complicated, 
so let us rename them.
```{r asm column rename, echo = TRUE}
ASM_data = ASM_data %>%
  rename(
    Geographic_Area_Name = Geographic.Area.Name..NAME.,
    NAICS_Code = X2017.NAICS.code..NAICS2017.,
    NAICS_Description = Meaning.of.NAICS.Code..NAICS2017_LABEL.,
    Year = Year..YEAR.,
    Total_Fringe_Benefits = Total.fringe.benefits...1.000...BENEFIT.,
    Fringe_Benefits_Error = Relative.standard.error.for.estimate.of.total.fringe.benefits......BENEFIT_S.,
    Total_Materials_Cost = Total.cost.of.materials...1.000...CSTMTOT.,
    Materials_Cost_Error = Relative.standard.error.for.estimate.of.total.cost.of.materials.......CSTMTOT_S.,
    Cost_Materials_Used = Cost.of.materials..packaging..etc..used...1.000...CSTMPRT.,
    Materials_Used_Error = Relative.standard.error.for.estimate.of.cost.of.materials..packaging..etc..used.....CSTMPRT_S.,
    Cost_Resales = Cost.of.resales...1.000...CSTRSL.,
    Resales_Error = Relative.standard.error.for.estimate.of.cost.of.resales......CSTRSL_S.,
    Cost_Fuels_Consumed = Cost.of.purchased.fuels.consumed...1.000...CSTFU.,
    Fuels_Consumed_Error = Relative.standard.error.for.estimate.of.cost.of.purchased.fuels.consumed......CSTFU_S.,
    Cost_Electricity = Cost.of.purchased.electricity...1.000...CSTELEC.,
    Electricity_Error = Relative.standard.error.for.estimate.of.cost.of.purchased.electricity......CSTELEC_S.,
    Cost_Contract_Work = Cost.of.contract.work...1.000...CSTCNT.,
    Contract_Work_Error = Relative.standard.error.for.estimate.of.cost.of.contract.work......CSTCNT_S.,
    Electricity_Purchased_kWh = Quantity.of.electricity.purchased.for.heat.and.power..1.000.kWh...ELECPCH.,
    Electricity_Purchased_Error = Relative.standard.error.for.estimate.of.quantity.of.electricity.purchased.for.heat.and.power......ELECPCH_S.,
    Electricity_Generated_kWh = Quantity.of.generated.electricity..1.000.kWh...ELECGEN.,
    Electricity_Generated_Error = Relative.standard.error.for.estimate.of.quantity.of.generated.electricity......ELECGEN_S.,
    Electricity_Sold_kWh = Quantity.of.electricity.sold.or.transferred..1.000.kWh...ELECSLD.,
    Electricity_Sold_Error = Relative.standard.error.for.estimate.of.quantity.of.electricity.sold.or.transferred......ELECSLD_S.,
    Inventories_Beginning_Year = Total.inventories..beginning.of.year...1.000...INVTOTB.,
    Inventories_Beginning_Error = Relative.standard.error.for.estimate.of.total.inventories..beginning.of.year......INVTOTB_S.,
    Finished_Goods_Beginning = Finished.goods.inventories..beginning.of.year...1.000...INVFINB.,
    Finished_Goods_Beginning_Error = Relative.standard.error.for.estimate.of.finished.goods.inventories..beginning.of.year......INVFINB_S.,
    Work_In_Process_Beginning = Work.in.process.inventories..beginning.of.year...1.000...INVWIPB.,
    Work_In_Process_Beginning_Error = Relative.standard.error.for.estimate.of.work.in.process.inventories..beginning.of.year......INVWIPB_S.,
    Materials_Supplies_Beginning = Materials.and.supplies.inventories..beginning.of.year...1.000...INVMATB.,
    Materials_Supplies_Beginning_Error = Relative.standard.error.for.estimate.of.materials.and.supplies.inventories..beginning.of.year......INVMATB_S.,
    Inventories_End_Year = Total.inventories..end.of.year...1.000...INVTOTE.,
    Inventories_End_Error = Relative.standard.error.for.estimate.of.total.inventories..end.of.year......INVTOTE_S.,
    Finished_Goods_End = Finished.goods.inventories..end.of.year...1.000...INVFINE.,
    Finished_Goods_End_Error = Relative.standard.error.for.estimate.of.finished.goods.inventories..end.of.year......INVFINE_S.,
    Work_In_Process_End = Work.in.process.inventories..end.of.year...1.000...INVWIPE.,
    Work_In_Process_End_Error = Relative.standard.error.for.estimate.of.work.in.process.inventories..end.of.year......INVWIPE_S.,
    Materials_Supplies_End = Materials.and.supplies.inventories..end.of.year...1.000...INVMATE.,
    Materials_Supplies_End_Error = Relative.standard.error.for.estimate.of.materials.and.supplies.inventories..end.of.year.......INVMATE_S.,
    Total_Capital_Expenditures = Total.capital.expenditures..new.and.used....1.000...CEXTOT.,
    Capital_Expenditures_Error = Relative.standard.error.for.estimate.of.total.capital.expenditures..new.and.used........CEXTOT_S.,
    Capital_Buildings = Capital.expenditures.for.buildings.and.other.structures...1.000...CEXBLD.,
    Capital_Buildings_Error = Relative.standard.error.for.estimate.of.capital.expenditures.for.buildings.and.other.structures......CEXBLD_S.,
    Capital_Machinery = Capital.expenditures.for.machinery.and.equipment...1.000...CEXMCH.,
    Capital_Machinery_Error = Relative.standard.error.for.estimate.of.capital.expenditures.for.machinery.and.equipment......CEXMCH_S.,
    Capital_Automobiles = Capital.expenditures.for.automobiles..trucks..etc..for.highway.use...1.000...CEXMCHA.,
    Capital_Automobiles_Error = Relative.standard.error.for.estimate.of.capital.expenditures.for.automobiles..trucks..etc..for.highway.use......CEXMCHA_S.,
    Capital_Computers = Capital.expenditures.for.computers.and.peripheral.data.processing.equipment...1.000...CEXMCHC.,
    Capital_Computers_Error = Relative.standard.error.for.estimate.of.capital.expenditures.for.computers.and.peripheral.data.processing.equipment.......CEXMCHC_S.,
    Capital_Other = Capital.expenditures.for.all.other.machinery.and.equipment...1.000...CEXMCHO.,
    Capital_Other_Error = Relative.standard.error.for.estimate.of.capital.expenditures.for.all.other.machinery.and.equipment......CEXMCHO_S.
  )
```

Let's take a look at the ASM data after renaming the variables to see if it
looks any better.
```{r asm second glimpse, echo = TRUE}
glimpse(ASM_data)
```

We can see that there are letters in some of the cells. Using the notes 
provided by the original dataset, let's replace those letters with descriptions.
```{r asm description replacements, echo = TRUE}
replacements = c(
  "D" = "Withheld",
  "N" = "Unavailable",
  "S" = "Poor Estimate",
  "X" = "Not Applicable",
  "A" = "High Error",
  "r" = "Revised",
  "s" = "Moderate Error"
)

ASM_data = ASM_data %>%
  mutate(across(everything(), ~ ifelse(. %in% names(replacements), replacements[.], .)))

glimpse(ASM_data)
```

There are some columns that we will not need for our analysis at all, so let's 
remove them to make the dataset easier to work with.
```{r asm removing unwanted variables, echo = TRUE}
ASM_data = ASM_data %>%
  select(-c(
    Cost_Fuels_Consumed,
    Fuels_Consumed_Error,
    Cost_Electricity,
    Electricity_Error,
    Cost_Contract_Work,
    Contract_Work_Error,
    Electricity_Purchased_kWh,
    Electricity_Purchased_Error,
    Electricity_Generated_kWh,
    Electricity_Generated_Error,
    Electricity_Sold_kWh,
    Electricity_Sold_Error
  ))

colnames(ASM_data)
```
We are also going to drop the NAICS Code column because the description already
tells us the industry so that row is redundant. 
```{r asm remove naics code, echo = TRUE}
ASM_data = ASM_data %>% select(-NAICS_Code)
```

We are going to split the dataset and keep the relative standard errors all 
together and replace the symbols with NA so we can do analysis.
```{r asm split relative error, echo = TRUE}
asm_relative_error = ASM_data %>% select(ends_with("Error"))
glimpse(asm_relative_error)
asm_relative_error = asm_relative_error %>% mutate(across(where(is.character), 
                          ~ case_when( . %in% c("Withheld","Unavailable",
"Poor Estimate","Not Applicable", "High Error","Revised", "Moderate Error") ~ NA_real_, 
  TRUE ~ as.numeric(.))))
```


Checking how many NA's are in the dataset. 
```{r asm na check 3?, echo = TRUE}
asm_relative_error %>% summarize_all(~ sum(is.na(.)))
```

Now looking at the ASM data we see the numerical columns are character we want 
to change that. 
1. We are going to remove the relative standard error subscript 
2. Then mutate across and parse the numbers 
3. There are warnings because some of the rows had characters in it 
(specifically withheld) so those values are just replaced by NA's which is fine.

```{r ASM columns from character to numeric, echo = TRUE}
ASM_data = ASM_data %>% select(!(ends_with("Error")))
ASM_data = ASM_data %>% mutate(across(.cols = -c(1,2), 
                                      ~ . %>% str_remove("(s)$") %>% parse_number()))
glimpse(ASM_data)
```
Checking how many NA's are in the "second dataset"

```{r ASM na check 4, echo = TRUE}
ASM_data %>% summarize_all(~ sum(is.na(.)))
```
Writing the data as a csv and storing it as a csv 
```{r asm csvs, echo = TRUE}
write_csv(ASM_data, "asm_data.csv")
write_csv(asm_relative_error, "asm_relative_error.csv")
```

